---
title: "Practice writing R functions"
subtitle: "That are relevant to simulation studies"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

# General structure of a function

Note that this is pseudo-code only: chunk is set not to run (`eval=FALSE`).

#basics of functions

data <- "my string"

whatever_function_one <- function(data) {
  
  #dependencies
  require(dplyr)
  
  # tests/checks
  if(is.data.frame(data)){
    stop("argument 'data' must be a data frame or tibble")
  }
  
  # do stuff
  output <- data |> 
    mutate(var=var + 1) # lets assume var exists
  
  # do stuff
  
  return(output)
  
}


```{r eval=FALSE}

# define function
function_name <- function(argument_1, # first arugment is often the data, if the function takes a data frame as an argument
                          argument_2 = "default", # arguments can have defaults
                          argument_3) {
  # required packages
  require(dplyr)
  
  # checks
  # well written functions contain checks. 
  # e.g., if the function assumes that argument_1 is a data frame, check that this is the case.
  # note that it is more useful to write the function first and add checks later.
  if(!is.data.frame(argument_1)){
    stop("argument_1 must be a data frame")
  }
  
  # code that does things
  object_to_be_returned <- input_data_frame |>
    # do things
    mutate(value = value + 1)
  
  # object to be returned
  return(object_to_be_returned)
}

# use function
function_name()

```

# Ideas for useful functions

Many of these functions have already been implemented by existing R packages (e.g., packages in the {[easystats](https://easystats.github.io/)} universe). However, our goal is not merely to find an existing solution, but to write one ourselves for practice.

## Relevant to generating data

# HOW DOES ANY OF THE FOLLOWING INFLUENCE COHEN'S D IN ANYWAY 

- Generate data from a uniform distribution and return a data frame.
- Generate bounded data, e.g., responses that are continuous but must be between 1 and 10. # a good question would be how bounding influences cohen's D
- Generate Likert data, e.g., responses that are whole numbers between 1 and 7, but which follow an underlying normal distribution.
- Simulate outliers, e.g., from careless responding or bots. 
- Generate item-level data for cross sectional studies [useful but too complicated]

## Relevant to analyzing data

- Convert a *p* value's significance to create APA-format table stars (i.e., "***" vs. "**" vs. "*" vs "ns")

```{r}

p_con <-function(p_value){
  
  #check results from data frame and convert

  if(p_value < .001){
    new_p <- "***"
  } else if(p_value < .01){
    new_p <- "**"
  } else if (p_value < .05){
    new_p <- "*"
  } else if(p_value >=.05){
    new_p <- "ns"
  }
  
    # return
  return(new_p) 
}

p_value <- 0.01

p_con(p_value)
```


- Convert a Cohen's *d* estimate to an interpretation
- Fit a correlation test and extract the p value and correlation
- Fit a regression and extract key results (p values, Beta estimates, etc)
- Fit and extract Cronbach's alpha [requires item level data]
- Choose another kind of test, fit it and extract its key information (eg estimate, p value), such as an assumption test, so that we could simulate the utility of tests of assumptions.
- Simulate publication bias by labeling a given study as "published" or "unpublished" based on a combination of its *p* value and a defined probability of (non)significant studies being published or not. 

## Relevant to summarizing simation results across iterations

- Summarize a column of data into a string that summarizes its mean and SD, which could be pasted directly into a manuscript. i.e., taking the form "M = XX.X (SD = XX.X)", with rounding and retention of lagging zeros.
- A function that rounds all numeric variables in a data frame by a given number of places.

# Session info

```{r}

sessionInfo()

```


